---
title: "Owens Valley Vegetation Condition"
subtitle: "2021 Annual Report"
author: "Water Department - County of Inyo"
date: "11/8/2021"
output: 
 html_document: 
   code_folding: hide
   fig_caption: yes
   toc: yes
   toc_depth: 5
   toc_float:
     collapsed: no
     smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE, message = FALSE, warning = FALSE}
library(targets)
library(tidyverse)
library(rmarkdown)
library(DT)
library(htmlwidgets)
library(sf)
library(tmap)
library(tmaptools)
# library(here)
library(ggpmisc)
library(ggpubr)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(tidyverse.quiet = TRUE)
```


```{r tar-load-initial}
tar_load(n_parcels_sampled)
tar_load(cYear)
tar_load(n_transects_sampled)
```


# Background
The Inyo County Water Department monitors groundwater-dependent ecosystems potentially affected by groundwater pumping. Data collection and statistical protocols are described in the Technical appendix to the Water Agreement (Green Book Box I.C.1.a.ii, [revised 2017](https://www.inyowater.org/wp-content/uploads/2017/09/GBMemo_SCfeb2017.pdf)). The primary purpose of this monitoring is to detect any “Significant decreases and changes in Owens Valley vegetation from conditions documented in 1984 to 1987". Vegetation management goals of the Agreement are based on canopy cover and species composition recorded during LADWP’s 1984-87 baseline vegetation parcel mapping and vegetation sampling effort.  To evaluate the condition of the vegetation, field crews of ICWD and LADWP monitor vegetation at permanent locations within a subset of the groundwater-dependent parcels potentially affected by pumping.

The purpose of this report is to provide the Inyo-LA Technical Group with an annual summary of the groundwater-dependent vegetation parcels on LA-owned land that show measurably different vegetation characteristics from the 1984-87 vegetation baseline. 

Green Book Box I.C.1.a.ii., "Transects for monitoring vegetation response to pumping", was revised by the Technical Group and Standing Committee in [2017](https://www.inyowater.org/wp-content/uploads/2017/09/GBMemo_SCfeb2017.pdf) to reflect agreement on how to jointly manage the vegetation monitoring program, including management of a shared database and which statistical procedures and parameters would be used to flag a measurable deviation from baseline conditions within vegetation parcel(s). Any parcel or group of parcels exhibiting a persistent deviation from baseline determined by these statistical procedures is then evaluated using the procedures in Green Book Section I.C.1.b. Determining Attributability. 

This research website provides this annual measurability analysis of Green Book Section I.C.1.a. The procedures described in Green Book Box I.C.1.a.ii. are translated here into a data pipeline using the `targets` R package, so that the implementation of the procedures are transparent, reproducible and accessible. The `workflowr` R package is used to create reproducibility checks and a version controlled github repository including this research website. The github repository can be accessed by clicking the `source` icon at the top right of this page.

To reproduce the results, the github repository can be cloned or downloaded, the `_targets.R` script is run in Rstudio which sources the `functions.R` script. Once the targets are created, the `report.rmd` is run to reproduce the `report.html` (this webpage).


Links to governing documents (hosted at [inyowater.org](https://www.inyowater.org/))

Inyo/LA Long Term Water Agreement([LTWA](https://www.inyowater.org/documents/governing-documents/water-agreement/))

LTWA Technical Appendix ([Green Book](https://www.inyowater.org/documents/governing-documents/green-book/))

Green Book Box I.C.1.a.ii.[TRANSECTS FOR MONITORING VEGETATION RESPONSE TO PUMPING](https://www.inyowater.org/wp-content/uploads/2017/09/GBMemo_SCfeb2017.pdf)

---

Green Book Section I.C, Impact Determination and Mitigation

"Among the primary goals of the Agreement are to manage groundwater pumping and surface water management practices as follows: 1) to avoid causing significant decreases in live vegetation cover; 2) to avoid changing a significant amount of vegetation from one classification to a lower (alphabetically) classification; 3) to avoid causing other significant effects on the environment; and 4) in a manner consistent with State and Federal laws pertaining to rare and endangered species. If any of these goals are not achieved, feasible mitigation of the affected area will be implemented. However, mitigation is not considered a primary management tool, but rather a secondary tool that will be employed should impacts occur that are inconsistent with the goals of the Agreement."

"This section outlines a procedure for determining whether decreases and/or changes in vegetation or other significant effects on the environment have occurred or are occurring in a given management area. It describes the process the Technical Group will follow to ascertain whether a change is significant, and thus, whether it requires mitigation. It also describes how the Technical Group will develop and implement a mitigation plan and monitoring and reporting program."


Section I.C.1, Determination of Significant Impacts

"A significant decrease or change or other significant effect on the environment will be mitigated if it is measurable, attributable to groundwater pumping or surface water management practices, and significant. The Agreement provides that the determination of significance of an impact, and thus, whether it must be mitigated, will be made on a case by case basis. The steps in the case by case analysis are described below."

Section I.C.1.a (Determining Measurability)

"In determining whether a change in vegetation cover or composition is measurable, the Technical Group will consider all relevant factors, including:

Comparison of current vegetation cover and composition in the affected area with similar data taken during other time periods, including the 1984-87 vegetation inventory data.

Comparison of vegetation cover and composition at the affected area with vegetation data from one or more control sites located in areas which have similar vegetation, soil, and precipitation conditions.

Comparison of the ratio of recently deceased vegetation to live vegetation in the affected area with other areas not affected by pumping and with similar vegetation cover and composition, soil, and precipitation conditions.

Use of air photo and remote sensing techniques to assist in making comparisons of conditions during different time periods and in mapping the affected area.

Comparison of data from randomly selected transects with similar data taken during other time periods, including the 1984-87 vegetation inventory data. This method will be employed in areas where monitoring site data does not exist, or where data covers an insufficient time period. Such transects will be performed as described in Box I.C.1.a.ii."

The updated Green Book BOX I.C.I.a.ii. outlines a set of procedures to test measurability based on the transect-level monitoring data's statistical deviation from baseline transect-level data.

Section I.C.1 describes

"A determination of measurability will be made if *any* of the relevant factors considered indicate even a *small* documentable change in vegetation cover or composition has occurred."

# Methods

### Baseline
From September 1984 to Nov 1987, LADWP inventoried and mapped vegetation into 2126 polygons of similar vegetation type, ‘vegetation parcels’ (223,168 acres).  Most of these lands were characterized as nonphreatophytic plant communities (Green Book management type A). The Green Book vegetation monitoring program is focused on groundwater-dependent parcels (Green Book types B, C, D and some E) and primarily those potentially affected by groundwater pumping. 

### Green Book Box I.C.1.a.ii{data-short-title="Green Book Methods"}

[**TRANSECTS FOR MONITORING VEGETATION RESPONSE TO PUMPING**](https://www.inyowater.org/wp-content/uploads/2017/09/GBMemo_SCfeb2017.pdf)

Vegetation transects are included within the Green Book to serve two purposes: 1) to estimate transpiration from a monitoring site, and 2) for use in determining whether vegetation has decreased or changed significantly from the previous cover. 

1) Detailed measurements of leaf area index shall be made at each of the monitoring sites using the techniques described in Section III.C. These measurements will be used to estimate evapotranspiration from the vegetation at the monitoring site for comparison to available soil water and, ultimately, to project plant-soil water balance and the need for water table recovery. Vegetation transects shall also be used in cases of suspected vegetation changes due to groundwater pumping. However, rather than using the intensive sampling technique of Section III.D for calculating evapotranspiration, plant cover shall be measured by the line-point technique described below.
 
2) During the 1984-87 inventory, most parcels were sampled with line-point transects of 100 feet in length, with sampling points at one-foot intervals, providing a two-dimensional representation of vegetation within the parcel. At each one-foot marker, the first contact with the uppermost layer of live plant cover was recorded. Cover and species composition were calculated from all sampling points along the transect.

The 1984-87 inventory shall be used as a "baseline" to determine whether vegetation cover and/or species composition has changed. This inventory is the only one of sufficient accuracy to permit comparison of species composition and cover.

A subset of the vegetation parcels mapped during the 1984-87 baseline inventory shall be annually monitored to accommodate statistical comparison with data collected during the baseline inventory. Such monitoring may rely on repeat measurement of georeferenced locations using the line-point-intercept method to track live cover and species composition over time. The baseline inventory was summarized at the parcel scale, thus for statistical comparison, transect locations derived from a set of random locations and azimuths provide a valid statistical comparison. 

Parcels to be monitored were initially selected based on meeting one or more of the following criteria: (1) parcel contained a permanent monitoring site; (2) baseline inventory data were collected for the parcel; (3) parcel was in close proximity to a pumping well; (4) information of past and current land use for parcel was available; (5) parcel was representative of one of the groundwater-dependent plant communities originally mapped during the baseline inventory; (6) soil characterization was available for the parcel; (7) characterization of the landscape position was available for the parcel. In 2015, Inyo County and Los Angeles began a joint monitoring program utilizing a combination of parcels and transects evaluated and agreed upon by the staffs of both parties. As of 2016, 1,688 transects across 141 parcels are jointly monitored by both parties.   

Approximately 100 of the 141 parcels monitored in 2016 will be jointly selected and monitored each year by Inyo County and Los Angeles. Prior to the field season, staff from both parties will determine which parcels of the 141 will be monitored. Existing permanent transects will be used for those parcels selected unless the staffs of both parties agree that a change in transect location is justified and necessary. Transects will be evenly split between ICWD and LADWP. Either party may independently conduct monitoring of additional parcels or transects; however, any data so collected shall be provided to the other party. 

The field protocol and calibration for field observations for the joint line-point monitoring program largely follows Herrick et al.(2016). At the onset of the monitoring season, staff from both parties shall meet in the field to discuss methods and to calibrate all field personnel.  In summary the field protocol is as follows: 

1. navigate to the transect start point with a handheld GPS device; 
2. verify presence of an aluminum tag which has been previously attached to a nearby object (i.e. shrub, debris, etc.); 
3. use a photo taken from the previous year to precisely locate the transect start point and to examine the accuracy of the azimuth used during the previous year; 
4. if the start point on the photo is off by more than 6 meters from the point indicated by a GPS reading, or a start point cannot be positively identified, a start point will be placed at the point indicated by the GPS device;
5. if the designated transect azimuth is off by more than ±5 degrees from the azimuth depicted within the previous year’s transect photo, the designated transect azimuth will be used. If not, the azimuth depicted within the previous year’s photo will be upheld;
6. stretch a tape measure to the direction of the pre-established azimuth;
7. for each transect, notes shall be taken on whether an aluminum tag is present, and whether or not the current year’s start point and compass azimuth matches those depicted within the previous year’s transect photo; 
8. record species identity at each half meter starting at 0.5 m and ending at 50 m yielding 100 possible hits; 
9. place a dry erase board with the parcel name, transect number, azimuth and date at the base on the transect line and take a picture from the start point toward the end point that captures the entire transect.

All live cover is tallied for each species for each transect. Data are exchanged at the end of the field season after each party digitally enters and proofs field data. 

The field technique and sampling design may be modified to permit detailed statistical comparison if deemed necessary in the future. Statistical analysis will be used to determine the measurability (statistical significance) of vegetation changes from the 1984-87 baseline inventory maps. Such an analysis may include, but is not limited to, Welch’s test (t-test with unequal variance), Wilcox test (Mann-Whitney test), Analysis of Variance (ANOVA) and Generalized Linear Model (GLM) for vegetation cover and permutational Multivariate Analysis of Variance (PERMANOVA) and Nonmetric Multidimensional Scaling (NMDS) for vegetation composition. For parcels with small baseline inventory sample sizes (e.g. 1-4 transects), a one-sample t-test may be performed using the baseline inventory sample mean as the null hypothesis for the test. Determination of measurability will be made by the Technical Group on case by case basis in accordance with Water Agreement IV.B and Green Book Section I.C.

Box I.C.1.a.ii shall be modified as necessary in the future to incorporate new or modified field techniques, technology, and/or analytical methods.  Such changes will be jointly developed by the staffs from both Inyo County and Los Angeles and will be presented to the Technical Group for consideration.


 
### Parcel Selection
Parcels were initially selected based on meeting one or more of the following criteria: (1) the parcel contained a permanent monitoring site where soil/vegetation water balance is calculated; (2) baseline data was collected for the parcel; (3) the parcel was in close proximity to a pumping well; (4) information of past and current land use for the parcel was available.  

<iframe src="https://inyocounty.maps.arcgis.com/apps/mapviewer/index.html?webmap=078c46d1d1714b58abbc79aa0d2066cc" height="405" width="720" style="border: 1px solid #464646;" allowfullscreen="" allow="autoplay" data-external="1"></iframe>



### Control Wellfield Designation
Parcels were classified as either belonging to a wellfield group or control group based on criteria derived from groundwater drawdown during the period of maximum pumping rate that occurred between 1987 and 1993. Parcels were assigned to the wellfield group if (1) kriged DTW estimates exceeded 1-m water-table drawdown during 1987-1993 or (2) they were located at sites corresponding to modeled drawdown contours greater than 10 ft. Parcels were assigned to the control group if (1) kriged DTW estimates were less than 1 m and (2) they were located at sites corresponding to modeled drawdown contours less than 10 ft. If the kriged DTW estimates were not reliable owing an inadequate test-well network near the vegetation parcel, then the groundwater-flow model estimate of the 10-ft drawdown contour was used as the sole criteria to designate parcels as either wellfield or control. An exception to the above criteria was applied to parcels associated with drawdown contours greater than 10-ft yet located near a surface water source (specifically, a canal, sewer pond, creek, river, or a ground water seepage source) that would lessen local drawdown effects—these parcels were classified as control. Some parcels currently in the wellfield group have higher water tables than during the 1987 to 1993 period, but remain in the wellfield group because of their close proximity to pumping wells and potential for pumping-induced drawdown.  Each parcel is also classified by its Green Book management type, Holland plant community type and by its status as either wellfield or control.

### Sample size
Most parcels were sampled in 1984-1987 using line-point-intercept sampling. Some parcels were not directly sampled but rather assigned cover and composition values from parcels with similar vegetation conditions. Welch’s t-test for unequal variance was used to evaluate in which parcels and in which year(s) total perennial cover and perennial grass cover has significantly differed from baseline. Since the sample standard deviation is used to construct 95% confidence interval, this method can be used for parcels in which baseline data contained more than one transect or nonzero sample variance. Where unmeasured parcels during baseline subsequently inherited a single baseline value from nearby parcels, a one-sample t-test is used to determine significance from baseline. For parcels with low baseline sample sizes, procedures in Green Book Box I.C.1.a.ii were followed: "For parcels with small baseline inventory sample sizes (e.g. 1-4 transects), a one-sample t-test may be performed using the baseline inventory sample mean as the null hypothesis for the test".

### Data

**Field Data (line-point-intercept)**

The number of parcels sampled each year as well as the number of transects sampled per parcel has varied due to fluctuations in annual staffing.  Thus, some parcels have varying numbers of transects sampled across time.  Other parcels have not been sampled continuously during the entire monitoring period. In `r cYear`, `r n_parcels_sampled` were sampled.

Perennial species cover is considered in this report, because annual species are not dependent on groundwater.  Perennial cover was further aggregated into grass, non-gramminoid herbaceous (herb), and shrub.  In order to analyze the changes in the composition of total perennial cover, the proportion of shrub, herb and grass cover as a fraction of total perennial cover was calculated. Transect data are summarized for each parcel and year using the arithmetic average, creating a history of cover over time for each parcel.  

**Remote Sensing**

Normalized difference vegetation index (NDVI) derived from Landsat 5/7/8 was extracted from google earth engine using the python API and an open source toolbox [ee-tools](https://github.com/Map-Water/ee-tools). Pixels were zonally averaged to the parcel polygons over July 15-Sep 15, creating a full history of remotely-sensed vegetation change starting during the baseline period in 1985. The Landsat dataset is produced by NASA/USGS, and the Landsat Science Team does the processing (masking clouds, preparing best images for 8-day,16-day images) before it is archived to the dataset on Google Earth Engine.  

**Precipitation**

Precipitation dataset was acquired using the gridMET/METDATA dataset produced at the University of Idaho and provided by MapWater consultants. It is a daily dataset of historically observed meteorological variables from Jan 1, 1979 to 2-days lag from the current date. It is produced over the contiguous United States (CONUS). GridMet is produced by bias correcting the daily NLDAS2 dataset to monthly PRISM values producing values on a 4-km (1/24-deg) grid (climateengine.org). 

### `targets`
The `Targets` R package is a data pipeline toolkit allowing dependency awareness and automatic rerunning of code that is downstream from dependency changes. A single `_targets.R` script was constructed specifying which datasets are input into which functions, and which outputs are then input into further downstream functions. The functions to process and analyze the data are maintained in a single file from `code/R/functions.R` and the `_targets.R` script sources these functions to make the pipeline.

The targets package `tar_visnetwork()` function output shows the data pipeline in an interactive format. Links are dependencies between functions and targets (input files, intermediate data transformations, model objects, plots and tables etc) specifying which targets are upstream from others and which are out of date. Functions are shown as triangles and targets are shown as circles. Fill color indicates whether or not upstream dependencies are 'Up to date' (green) or 'Outdated' (blue). Only the downstream dependencies of outdated targets or functions will be re-evaluated when the pipeline is rerun using `tar_make()`.


```{r visnetwork}
tar_visnetwork()
```

View [`_targets.R`](https://github.com/zach-nelson/wflow-veg/blob/master/_targets.R) script file.

View [`functions.R`](https://github.com/zach-nelson/wflow-veg/blob/master/code/R/functions.R)


### `workflowr`
Once the targets are defined in the `_targets.R` script, running tar_make() stores the targets in the `_targets` subdirectory under the root of the project, where the .Rproj is located. Now any `.Rmd` anywhere within the project directory can access these targets that were made from the dependency-aware pipeline. The default `workflowr` project directory defaults `.Rmd` files to be located in the `analysis` directory and these `analysis/file.Rmd` can access the targests. The `workflowr::wflow_build()` will render the `.Rmd` to an html website in the `docs` directory that can be version controlled on github and hosted on github-pages. 

`workflowr` research websites include seven reproducibility checks including that the output was derived from a 'committed' file to version control, the R environment was empty, the seed was set, session information was recorded, no cache, file paths were relative, and which repository version created the website view.



```{r load-targets}
tar_load(n_parcels_all_years)
# tar_load(parcel_test_sums)
tar_load(parcel_test_sums)


tar_load(icwd_ladwp_bind)
tar_load(lpt_updated_master)
tar_load(transects)
tar_load(parcels)
tar_load(parcels_shp)
tar_load(parcels_deltas)#parcel-year-deltas
# tar_load(parcel_year_meta_2samp_results)
# tar_load(parcel_year_meta_1samp_results)
tar_load(parcel_year_meta_combined_results)
tar_load(deltas_ttest_att)
tar_load(wellcont_means)
tar_load(wellcont_means_rarefied)
tar_load(plot_wellcontrol)
tar_load(boxplot.w.c)
tar_load(trends.w.c)

tar_load(plot_1sample_timeseries)
tar_load(plot_2sample_timeseries)
tar_load(parcel_datatable)
tar_load(parcel_datatable_chronic)
tar_load(parcel_datatable_significant)
tar_load(attributes_reinv)
tar_load(attributes_pfix)
tar_load(panel_map_lw)
tar_load(panel_map_bp)
tar_load(panel_map_ta)
tar_load(panel_map_ts)
tar_load(panel_map_io)
tar_load(panel_map_ss)
tar_load(panel_map_bg)
# tar_load(panel_map_lp)
# tar_load(monsites_shp)

tar_load(parcel_year_meta_2samp_results_grass)
tar_load(parcel_year_meta_1samp_results_grass)
tar_load(parcel_year_meta_2samp_results)
tar_load(parcel_year_meta_1samp_results)
```

# Results

```{r parcel-attributes,include=FALSE,eval=FALSE}
# Parcel Attributes
attributes_reinv%>%
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 2, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )
```

```{r cyr-tran-species,include=FALSE, eval=FALSE}
# Current year data
icwd_ladwp_bind %>% filter(Cover > 0)%>%
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 2, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )
```

```{r tran-species,include=FALSE, eval=FALSE}
# Total combined data
lpt_updated_master %>% filter(Cover > 0)%>% 
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 2, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )

```

```{r tran-ftype,include=FALSE, eval=FALSE}
# Aggregated to transect cover type
transects%>% 
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 3, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )
```


## Wellfield Control Comparison{data-short-title="Wellfield-Control"}


For an overview of the pumping effect on GDEs in Owens Valley, parcels were classified as either belonging to a wellfield group or control group, based on drawdown from 1987 and 1993, the period of maximum pumping rate. ^[Parcels were assigned to the wellfield group if (1) kriged depth-to-water (DTW) estimates exceeded 1-m water-table drawdown during 1987-1993 or (2) they were located at sites corresponding to modeled drawdown contours greater than 10 ft. Parcels were assigned to the control group if (1) kriged DTW estimates were less than 1 m and (2) they were located at sites corresponding to modeled drawdown contours less than 10 ft. If the kriged DTW estimates were not reliable owing an inadequate test-well network near the vegetation parcel, then the groundwater-flow model estimate of the 10-ft drawdown contour was used as the sole criteria to designate parcels as either wellfield or control. An exception to the above criteria was applied to parcels associated with drawdown contours greater than 10-ft yet located near a surface water source (specifically, a canal, sewer pond, creek, river, or a ground water seepage source) that would lessen local drawdown effects — these parcels were classified as control.]

Here the subset of wellfield and control parcels that have been sampled in every year 1992-`r cYear` are considered to remove any trend potentially associated with group sample sizes that may have fluctuated over the years. Wellfield and control group averages are plotted for comparison over time and linear trends are reported for total cover and for the perennial herbaceous component and woody shrub component of total cover separately.   

```{r plot-wellcontrol}
plot_wellcontrol
```

```{r plot-wellcontrol-trends}
trends.w.c
```

```{r plot-wellcontrol-boxplot1}
boxplot.w.c
```

 
## Parcel Cover

In the `r cYear` growing season, ICWD and LADWP monitored `r n_parcels_sampled` vegetation parcels with approximately `r n_transects_sampled` vegetation transects using the line-point-intercept method described Green Book Box I.C.1.a.ii. 


```{r parcel-deltas-table}
parcels_deltas %>% 
  select(Parcel, Year,Cover, Cover.Delta, Grass, Grass.Delta, Shrub, Shrub.Delta) %>% 
  mutate(across(Cover:Shrub.Delta, round, 1)) %>% 
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Blfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 2, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )
# %>% formatRound(c('july','oct'),2)
# deltas
```

## Grass Cover Change

The baseline perennial vegetation cover and subset of grass cover statistically compared to `r cYear` data using either Welch Two Sample t-test or One Sample t-test - depending on the baseline sample size following Box I.C.1.a.ii.


```{r parcel-yr-ttest-cyear}
parcel_year_meta_combined_results %>% 
  filter(Year.x == cYear) %>% 
  select(Parcel,cover_type,significance,  estimate, estimate1, estimate2, method,n.x, n.y, statistic, p.value) %>%  
  mutate(across(estimate:estimate2, round, 0)) %>%
  mutate(across(statistic:p.value, round, 3)) %>% 
  datatable(extensions = c('Buttons','FixedColumns'),
            filter = c("top"),
            options = list(
              dom = 'Blfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
              scrollX = TRUE,
              fixedColumns = list(leftColumns = 2, rightColumns = 0),
              pageLength = 5, 
              autoWidth = TRUE)
            )


# ,filter = c("top"),options = list(
  # pageLength = 5, autoWidth = TRUE)) 


# est, est1, est2, stat, pval, par, clow, chigh, method, alt, sig
```

## Effect size and persistence

above pivot wide the effect size of parcel vegetation change (Cover.Delta, Grass.Delta, Shrub.Delta) from baseline to `r cYear`, including the `r cYear` proportion of grass (pGrass), shrub (pShrub), parcel average perennial total cover (Cover), and subtypes Grass and Shrub cover.
Cover_sig.counter, Grass_sig.counter, 
```{r parcel-delta-sig-table}
# tar_read(
  parcel_datatable_significant
  
```

## Wellfield Parcels Below Baseline

These wellfield parcels were below baseline perennial cover in `r cYear`. `Cover.Delta` is the change from baseline. `sig.counter` is the number of years the parcel has been below baseline in a row, `sum.sig` is the number of total years it was significantly below baseline out of `n` years sampled, and `max.run` is the longest consecutive period the parcel was below baseline.

```{r datatable-chronic}
parcel_datatable_chronic
```

## Laws
```{r lw-map}

panel_map_lw
# knitr::include_graphics(panel_map_lw, error = FALSE)


```

## Big Pine
```{r bp-map}
panel_map_bp
# knitr::include_graphics(panel_map_bp, error = FALSE)

```

## Taboose-Aberdeen
```{r ta-map}
panel_map_ta
# knitr::include_graphics(panel_map_ta, error = FALSE)

```

## Thibaut-Sawmill
```{r ts-map}
panel_map_ts
# knitr::include_graphics(panel_map_ts, error = FALSE)

```

## Independence-Oak
```{r io-map}
panel_map_io
# knitr::include_graphics(panel_map_io, error = FALSE)

```

## Symmes-Shepherd
```{r ss-map}
panel_map_ss
# knitr::include_graphics(panel_map_ss, error = FALSE)

```

## Bairs-George
```{r bg-map}
panel_map_bg
# knitr::include_graphics(panel_map_bg, error = FALSE)

```


```{r 2sample-graph,include=FALSE,eval=FALSE}
# Graph two-sample ttest, These graphs could be split out by wellfield
plot_2sample_timeseries
```

```{r 1sample-graph,include=FALSE,eval=FALSE}
# Graph one-sample ttest
plot_1sample_timeseries
```

# Discussion

"To determine if a significant effect has occurred or is occurring, the Technical Group first shall evaluate the measurability of the effect" (Green Book Section I.C.1.a). "A determination of measurability will be made if *any* of the relevant factors considered indicate even a *small* documentable change in vegetation cover or composition has occurred."

A determination of measurability has been automated here in code so that the application of the statistical methods in the revised Green Book Box I.C.1.a.ii. are transparent and reproducible.

The list of parcels below show a measurable reduction in either total perennial cover or total perennial grass cover in `r cYear` in addition to the five previous years. These parcels, exhibiting measurable and persistent changes identified in this report meet the criteria for the Technical Group to evaluate the cause(s) of the apparent changes following Green Book Section I.C.1.b.

```{r datatable-chronic2}
# limited to only sig.counter grass.counter > 5 years
parcel_datatable_chronic
```


